---
title: "4_Demographics"
author: "Meg Hutch"
date: "September 22, 2019"
output: html_document
---

Objective: This script processes demographic data from REDCap for our ICP Cohort


# Packages
```{r Load Packages}
library(dplyr)
library(pastecs)
```

# Import Demographics Data
```{r}
demo <- read.csv("raw_data/BWHMGHPupillometry20_DATA_2020-08-22_1531.csv")

## separate file for surgery information
surgery <- read.csv("raw_data/4.16.21 Craniectomy Partners Demographics copy.csv")
```

# Clean Up Data

```{r}
## separate surgery information from the REDCap data
demo <- demo %>% filter(!redcap_repeat_instrument == "surgeries")
demo$redcap_repeat_instrument <- NULL
demo$redcap_repeat_instance <- NULL

# Change record_id name
colnames(demo)[1] <- "Study_ID"

```


# Race
```{r}
# 0, White | 1, Black or African American | 2, Asian | 3, Native Hawaiian or Other Pacific Islander | 4, American Indian or Alaska Native | 999, Other/Unknown												

#Rename Race columns
demo <- demo %>% rename(White = race___0,
                        Black = race___1,
                        Asian = race___2, 
                        Nat.Hawaiian_Pac.Isl = race___3, 
                        Amer.Ind_Alaska.Nat = race___4, 
                        Other_Unknown_Race = race___999)

#Create a multi_race category
demo$Race_counts <- rowSums(demo[c(4:9)] == 1) #2 means multiple race 
demo$Multi_Race <- ifelse(demo$Race_counts > 1, 1, 0)
```

# Rename Diagnosis Columns
```{r}
# 0, IPH (Intraparenchymal Hemorrhage) | 1, SAH (Subarachnoid Hemorrhage) | 2, SDH (Subdural Hemorrhage) | 3, EDH (Epidural Hemorrhage) | 4, TBI (Traumatic Brain Injury) | 5, Anterior Circulation Stroke | 6, Posterior Circulation Stroke | 7, Status Epilepticus | 8, IVH (Intraventricular Hemorrhage) | 9, Tumor | 10, Infection | 11, Hydrocephalus | 12, Cerebral Sinus Venous Thrombosis | 999, Other

# Rename Diangoses columns
demo <- demo %>% rename(IPH_Dx = primary_diagnosis___0, 
                      SAH_Dx = primary_diagnosis___1, 
                      SDH_Dx = primary_diagnosis___2, 
                      EDH_Dx = primary_diagnosis___3, 
                      TBI_Dx = primary_diagnosis___4, 
                      AC.Stroke_Dx = primary_diagnosis___5, 
                      PC.Stroke_Dx = primary_diagnosis___6, 
                      Status.E_Dx = primary_diagnosis___7, 
                      IVH_Dx = primary_diagnosis___8, 
                      Tumor_Dx = primary_diagnosis___9, 
                      Infection_Dx = primary_diagnosis___10, 
                      Hydrocephalus_Dx = primary_diagnosis___11, 
                      CVST_Dx = primary_diagnosis___12,
                      Other_Dx = primary_diagnosis___999)

# Final Diagnosis: 0, Spontaneous IPH | 1, Aneurysmal SAH | 2, Stroke | 3, Parenchymal TBI | 4, Brain Tumor | 5, Isolated Hydrocephalus without IVH | 6, Seizure or Status | 7, Infection | 8, SDH | 9, EDH | 10, Non aneurysmal SAH | 11, CSVT | 12, Isolated IVH | 13, Hypoxic Ischemic Injury | 999, Other
demo <- demo %>% rename(other_desc_final = other_final)

demo <- demo %>% rename(Spon.IPH_final = final_dx___0, 
                      ASah_final = final_dx___1, 
                      stroke_final = final_dx___2, 
                      parenchymal_tbi_final = final_dx___3, 
                      brain_tumor_final = final_dx___4, 
                      isolated_hydro_final = final_dx___5, 
                      seizure_status_final = final_dx___6, 
                      infection_final = final_dx___7, 
                      sdh_final = final_dx___8, 
                      edh_final = final_dx___9, 
                      non_asah_final = final_dx___10,
                      cvst_final = final_dx___11,
                      isolated_ivh_final = final_dx___12,
                      hypoxic_ischemic_injury_final = final_dx___13,
                      other_final = final_dx___999)
```

# Rename pathology side varaibles
```{r}
# 1, ACA territory | 2, MCA territory | 3, PCA territory | 4, Cerebellar | 5, Midbrain | 6, Pons | 7, Medulla | 9, Thalamus | 10, Watershed | 11, Frontal Lobe | 12, Temporal Lobe | 13, Parietal Lobe | 14, Occipital Lobe | 15, Corpus Callosum | 16, Caudate | 17, Internal Capsule | 18, Basal Ganglia | 19, Insula | 20, Putamen | 21, Globus Pallidus | 22, Corona Radiata | 23, Periventricular white matter | 24, Subcortical white matter | 25, Cingulate Gyrus | 26, Scattered Bilateral Cerebral | 27, Spine | 28, Ventricles | 29, Hippocampus | 999, Other

demo <- demo %>% rename(ACA_Path = pathology_side___1, 
                        MCA_Path = pathology_side___2, 
                        PCA_Path = pathology_side___3, 
                        Cerebellar_Path = pathology_side___4, 
                        Midbrain_Path = pathology_side___5, 
                        Pons_Path = pathology_side___6, 
                        Medulla_Path = pathology_side___7, 
                        Thalamus_Path = pathology_side___9, 
                        Watershed_Path = pathology_side___10, 
                        Frontal_lobe_Path = pathology_side___11, 
                        Temporal_lobe_Path = pathology_side___12, 
                        Parietal_lobe_Path = pathology_side___13, 
                        Occipital_lobe_Path = pathology_side___14, 
                        Corpus_Callosum_Path = pathology_side___15, 
                        Caudate_Path = pathology_side___16, 
                        Internal_Capsule_Path = pathology_side___17, 
                        Basal_ganglia_Path = pathology_side___18, 
                        Insula_Path = pathology_side___19, 
                        Putamen_Path = pathology_side___20, 
                        Globus_Pallidus_Path = pathology_side___21, 
                        Corona_Radiata_Path = pathology_side___22, 
                        Periventricular_White_Matter_Path = pathology_side___23, 
                        Subcortical_White_Matter_Path = pathology_side___24, 
                        Cingulate_Gyrus_Path = pathology_side___25, 
                        Scattered_Bilateral_Cerebral_Path = pathology_side___26, 
                        Spine_Path = pathology_side___27, 
                        Ventricles_Path = pathology_side___28, 
                        Hippocampus_Path = pathology_side___29, 
                        Other_Path = pathology_side___999, 
                        Other_Desc_Path = other_pathology)

#Add suffix to indicate columns for pathology side - This indicates L, R, or Bilateral - Should be ACA to Hippocampus primary
colnames(demo)[87:111] <- paste(colnames(demo)[87:111], "Path_Side", sep = "_")

#Change back mca_complete_incomplte
colnames(demo)[89] <- "mca_complete_incomplete"

# Recode all pathsides
demo$aca_Path_Side <- recode(demo$aca_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$mca_Path_Side <- recode(demo$mca_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

# 1, Complete | 0, Incomplete | 888, Not specified | 999, Unclear
demo$mca_complete_incomplete <- recode(demo$mca_complete_incomplete, '1' = "Complete", '0' = "Incomplete", '888' = "Not Specified", '999' = "Unclear")

demo$pca_Path_Side <- recode(demo$pca_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$cerebellar_Path_Side <- recode(demo$cerebellar_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$midbrain_Path_Side <- recode(demo$midbrain_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$pons_Path_Side <- recode(demo$pons_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$medulla_Path_Side <- recode(demo$medulla_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$thalamus_Path_Side <- recode(demo$thalamus_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$frontal_Path_Side <- recode(demo$frontal_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$temporal_Path_Side <- recode(demo$temporal_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$parietal_Path_Side <- recode(demo$parietal_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$occipital_Path_Side <- recode(demo$occipital_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$corpus_callosum_Path_Side <- recode(demo$corpus_callosum_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$caudate_Path_Side <- recode(demo$caudate_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$internal_capsule_Path_Side <- recode(demo$internal_capsule_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$basal_ganigla_Path_Side <- recode(demo$basal_ganglia_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$insula_Path_Side <- recode(demo$insula_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$putamen_Path_Side <- recode(demo$putamen_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$globus_pallidus_Path_Side <- recode(demo$globus_pallidus_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$corona_radiata_Path_Side <- recode(demo$corona_radiata_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$periventricular_white_matter_Path_Side <- recode(demo$periventricular_white_matter_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$subcortical_white_matter_Path_Side <- recode(demo$subcortical_white_matter_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$cingulate_gyrus_Path_Side <- recode(demo$cingulate_gyrus_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$hippocampus_primary_Path_Side <- recode(demo$hippocampus_primary_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")
```

# Create level names for variables that are coded
```{r}
# gender: 0, Female | 1, Male | 999, Other/Unknown
demo$gender <- recode(demo$gender, '0' = "Female", '1' = "Male", '999' = "Unknown")

# ethnicity: 0, Not Hispanic or Latino | 1, Hispanic or Latino | 999, Unknown
demo$ethnicity <- recode(demo$ethnicity, '0' = "Not_Hispanic", '1' = "Hispanic", '999' = "Unknown")

# IPH type:1, Hemorrhagic conversion | 2, Spontaneous | 3, Traumatic | 4, Vascular | 5, Secondary to tumor | 999, Unspecified
demo$iph_type <- recode(demo$iph_type, '1' = "Hem_Conversion", '2' = "Spontaneous", '3' = "Traumatic", '4' = "Vascular", '5' = "Sec_to_Tumor", '999' = "Unspecified")

# Vascular etiology: 1, AVM | 2, Aneurysm | 999, Unspesified
demo$vascular_etiology <- recode(demo$vascular_etiology, '1' = "AVM", '2' = "Aneurysm", '999' = "Unspecified")

# ac IPH: 1, Yes | 0, No | 999, Indeterminate/Unclear
demo$ac_iph <- recode(demo$ac_iph, '0' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")

# SAH type: 1, Aneurysmal | 2, Non-aneurysmal | 999, Unspecified
demo$sah_type <- recode(demo$sah_type, '1' = "Aneurysmal", '2' = "Non-Aneurysmal", '999' = "Unspecified")

# ac_sah: 1, Yes | 0, No | 999, Indeterminate/Unclear
demo$ac_sah <- recode(demo$ac_sah, '0' = "No", '1' = "Yes", '999' = "Unspecified")

# sdh etiology: 1, Traumatic | 2, Spontaneous | 999, Unspecified
demo$sdh_etiology <- recode(demo$sdh_etiology, '1' = "Traumatic", '2' = "Spontaenous", '999' = "Unspecified")

# sdh acuity: 0, Non-acute (fall/injury > 48 hrs) | 1, Acute (fall/injury < 48 hrs) | 2, Acute on chronic | 999, Unspecified
demo$sdh_acuity <- recode(demo$sdh_acuity, '0' = "Non-acute (> 48 Hours)", '1' = "Acute (< 48 Hours)", '2' = "Acute on Chronic", '999' = "Unspecified")

# ac sdh: 1, Yes | 0, No | 999, Indeterminate/Unclear
demo$ac_sdh <- recode(demo$ac_sdh, '0' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")

# tbi location: 1, Diffuse | 2, Focal | 3, Multifocal | 999, Unspecified
demo$tbi_location <- recode(demo$tbi_location, '1' = "Diffuse", '2' = "Focal", '3' = "Multifocal", '999' = "Unspecified")

# focal tbi: 1, Blunt | 2, Penetrating | 999, Unspecified
demo$focal_tbi_type <- recode(demo$focal_tbi_type, '1' = "Blunt", '2' = "Penetrating", '999' = "Unspecified")

# ac tbi: 1, Yes | 0, No | 999, Indeterminate/Unclear
demo$ac_tbi <- recode(demo$ac_tbi, '0' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")

# mass effect: 1, Yes | 2, No | 999, Indeterminate
demo$mass_effect <- recode(demo$mass_effect, '2' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")

# mass effect side: 1, Left | 2, Right | 999, Indeterminate
demo$mass_effect_side <- recode(demo$mass_effect_side, '1' = "Left", '2' = "Right", '999' = "Indeterminate/Unclear")

# mls: 1, Yes | 0, No
demo$mls <- recode(demo$mls, '1' = "Yes", '0' = "No")

# herniation: 1, Yes | 0, No
demo$herniation <- recode(demo$herniation, '1' = "Yes", '0' = "No")

#primary_diagnosis_4: this is really for edh_ac: 1, Yes | 0, No | 999, Indeterminate/ Unclear
demo$primary_diagnosis_4 <- recode(demo$primary_diagnosis_4, '0' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")

# change column name
demo <- demo %>% rename("ac_edh" = "primary_diagnosis_4")
```

# Herniation Type
```{r}
# herniation type: 1, Subfalcine | 2, Uncal | 3, Transtentorial | 4, Upward Cerebellar | 5, Tonsillar | 6, Upward Transtentorial | 999, Unspecified
demo <- demo %>% rename(Herniation_Subfalcine = herniation_type___1, 
                        Herniation_Uncal = herniation_type___2, 
                        Herniation_TransTentorial = herniation_type___3, 
                        Herniation_Up_Cerebellar = herniation_type___4, 
                        Herniation_Tonsillar = herniation_type___5, 
                        Herniation_Up_Transtentorial = herniation_type___6, 
                        Herniation_Unspecified = herniation_type___999) 
```

# Secondary Diagnosis
```{r}
# 0, IPH (Intraparenchymal Hemorrhage) | 1, SAH (Subarachnoid Hemorrhage) | 2, SDH (Subdural Hemorrhage) | 3, EDH (Epidural Hemorrhage) | 4, TBI (Traumatic Brain Injury) | 5, Anterior Circulation Stroke | 6, Posterior Circulation Stroke | 7, Status Epilepticus | 8, IVH (Intraventricular Hemorrhage) | 9, Tumor | 10, Infection | 11, Cerebral Edema | 12, Hydrocephalus | 13, Cerebral Sinus Venous Thrombosis | 999, Other
demo <- demo %>% rename(IPH_Dx2 = secondary_diagnosis___0, 
                      SAH_Dx2 = secondary_diagnosis___1, 
                      SDH_Dx2 = secondary_diagnosis___2, 
                      EDH_Dx2 = secondary_diagnosis___3, 
                      TBI_Dx2 = secondary_diagnosis___4, 
                      AC.Stroke_Dx2 = secondary_diagnosis___5, 
                      PC.Stroke_Dx2 = secondary_diagnosis___6, 
                      Status.E_Dx2 = secondary_diagnosis___7, 
                      IVH_Dx2 = secondary_diagnosis___8, 
                      Tumor_Dx2 = secondary_diagnosis___9, 
                      Infection_Dx2 = secondary_diagnosis___10, 
                      Cerebral_Edema_Dx2 = secondary_diagnosis___11, 
                      Hydrocephalus_Dx2 = secondary_diagnosis___12,
                      CVST_Dx2 = secondary_diagnosis___13,
                      Other_Dx2 = secondary_diagnosis___999)

# Notes
demo <- demo %>% rename(other_desc_dx2 = other_secondary_diagnosis)
```

# Code additional information - secondary diagnosis
```{r}
#iph_type2: 1, Hemorrhagic conversion | 2, Spontaneous | 3, Traumatic | 4, Vascular | 5, Secondary to tumor | 999, Unspecified
demo$iph_type2 <- recode(demo$iph_type2, '1' = "Hem_Conversion", '2' = "Spontaneous", '3' = "Traumatic", '4' = "Vascular", '5' = "Sec_to_Tumor", '999' = "Unspecified")

#vascular_etiology_secondary: 1, AVM | 2, Aneurysm | 999, Unspecified
demo$vascular_etiology_secondary <- recode(demo$vascular_etiology_secondary, '1' = "AVM", '2' = "Aneurysm", '999' = "Unspecified")

#ac_iph2: 1, Yes | 0, No | 999, Indeterminate
demo$ac_iph2 <- recode(demo$ac_iph2, '0' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")

#sah_type2: 1, Aneurysmal | 2, Non-aneurysmal | 999, Unspecified
demo$sah_type2 <- recode(demo$sah_type2, '1' = "Aneurysmal", '2' = "Non-Aneurysmal", '999' = "Unspecified")

#ac_sah2: 1, Yes | 0, No | 999, Indeterminate
demo$ac_sah2 <- recode(demo$ac_sah2, '0' = "No", '1' = "Yes", '999' = "Unspecified")

#sdh_etiology2: 1, Traumatic | 2, Spontaneous | 999, Unspecified
demo$sdh_etiology2 <- recode(demo$sdh_etiology2, '1' = "Traumatic", '2' = "Spontaenous", '999' = "Unspecified")

#blood_thinners2: 1, Yes | 0, No | 999, Unspecified 
demo$blood_thinners2 <- recode(demo$blood_thinners2, '0' = "No", '1' = "Yes", '999' = "Unspecified")

#sdh_acuity2: 0, Non-acute (fall/injury > 48 hrs) | 1, Acute (fall/injury < 48 hrs) | 999, Unspecified
demo$sdh_acuity2 <- recode(demo$sdh_acuity2, '0' = "Non-acute (> 48 Hours)", '1' = "Acute (< 48 Hours)", '999' = "Unspecified") 

#ac_sdh2: 1, Yes | 0, No | 999, Indeterminate
demo$ac_sdh2 <- recode(demo$ac_sdh2, '0' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")

## 12.1.2019: TBI_Locations aren't running since this dataset does not have data for these variables
## 01.09.2020: Also did not run; no data
## 01.24.2020: Also did not run; no data
## 08.22.2020: Also did not run; no data
#tbi_location2: 1, Diffuse | 2, Focal | 999, Unspecified
demo$tbi_location2 <- recode(demo$tbi_location2, '1' = "Diffuse", '2' = "Focal", '999' = "Unspecified")

#focal_tbi_type2: 1, Blunt | 2, Penetrating | 999, Unspecified
demo$focal_tbi_type2 <- recode(demo$focal_tbi_type2, '1' = "Blunt", '2' = "Penetrating", '999' = "Unspecified")

#ac_tbi2: 1, Yes | 0, No | 999, Indeterminate
demo$ac_tbi2 <- recode(demo$ac_tbi2, '0' = "No", '1' = "Yes", '999' = "Indeterminate/Unclear")
```

# Secondary Pathsides
```{r}
#1, ACA territory | 2, MCA territory | 3, PCA territory | 4, Cerebellar | 5, Midbrain | 6, Pons | 7, Medulla | 9, Thalamus | 10, Watershed | 11, Frontal Lobe | 12, Temporal Lobe | 13, Parietal Lobe | 14, Occipital Lobe | 15, Corpus Callosum | 16, Caudate | 17, Internal Capsule | 18, Basal Ganglia | 19, Insula | 20, Putamen | 21, Globus pallidus | 22, Corona Radiata | 23, Periventricular white matter | 24, Subcortical white matter | 25, Cingulate Gyrus | 26, Scattered Bilateral Cerebral | 27, Spine | 28, Ventricles | 29, Hippocampus | 999, Other

demo <- demo %>% rename(ACA_Path2 = secondary_pathside___1, 
                        MCA_Path2 = secondary_pathside___2, 
                        PCA_Path2 = secondary_pathside___3, 
                        Cerebellar_Path2 = secondary_pathside___4, 
                        Midbrain_Path2 = secondary_pathside___5, 
                        Pons_Path2 = secondary_pathside___6, 
                        Medulla_Path2 = secondary_pathside___7, 
                        Thalamus_Path2 = secondary_pathside___9, 
                        Watershed_Path2 = secondary_pathside___10, 
                        Frontal_lobe_Path2 = secondary_pathside___11, 
                        Temporal_lobe_Path2 = secondary_pathside___12, 
                        Parietal_lobe_Path2 = secondary_pathside___13, 
                        Occipital_lobe_Path2 = secondary_pathside___14, 
                        Corpus_Callosum_Path2 = secondary_pathside___15, 
                        Caudate_Path2 = secondary_pathside___16, 
                        Internal_Capsule_Path2 = secondary_pathside___17, 
                        basal_ganglia_Path2 = secondary_pathside___18, 
                        Insula_Path2 = secondary_pathside___19, 
                        Putamen_Path2 = secondary_pathside___20, 
                        Globus_Pallidus_Path2 = secondary_pathside___21, 
                        Corona_Radiata_Path2 = secondary_pathside___22, 
                        Periventricular_White_Matter__Path2 = secondary_pathside___23, 
                        Subcortical_White_Matter_Path2 = secondary_pathside___24, 
                        Cingulate_Gyrus_Path2 = secondary_pathside___25, 
                        Scattered_Bilateral_Cerebral_Path2 = secondary_pathside___26, 
                        Spine_Path2 = secondary_pathside___27, 
                        Ventricles_Path2 = secondary_pathside___28, 
                        Hippocampus_Path2 = secondary_pathside___29, 
                        Other_Path2 = secondary_pathside___999)

#Add suffix to indicate columns for pathology side - This indicates L, R, or Bilateral - Should be ACA2 to Hippocampus_secondary - should start at aca2 to hippocampus_secondary
colnames(demo)[187:211] <- paste(colnames(demo)[187:211], "Path_Side", sep = "_")

# Other pathside notes
demo <- demo %>% rename(other_desc_pathside2 = other_pathside)

#Change back mca_complete_incomplte
colnames(demo)[189] <- "mca_complete_incomplete2"

# Recode all pathsides
demo$aca2_Path_Side <- recode(demo$aca2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$mca2_Path_Side <- recode(demo$mca2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

# 12.1.2019; Did not run because no data here
# 01.09.2020; Also did not run because no data
# 01.24.2020; Also did not run because no data
# 08.22.2020; Also did not run because no data
# 1, Complete | 0, Incomplete | 888, Not specified | 999, Unclear
demo$mca_complete_incomplete2 <- recode(demo$mca_complete_incomplete2, '1' = "Complete", '0' = "Incomplete", '888' = "Not Specified", '999' = "Unclear")

demo$pca2_Path_Side <- recode(demo$pca2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$cerebellar2_Path_Side <- recode(demo$cerebellar2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$midbrain2_Path_Side <- recode(demo$midbrain2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$pons2_Path_Side <- recode(demo$pons2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$medulla2_Path_Side <- recode(demo$medulla2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$thalamus2_Path_Side <- recode(demo$thalamus2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$frontal2_Path_Side <- recode(demo$frontal2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$temporal2_Path_Side <- recode(demo$temporal2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$parietal2_Path_Side <- recode(demo$parietal2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$occipital2_Path_Side <- recode(demo$occipital2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$corpus2_callosum_Path_Side <- recode(demo$corpus_callosum2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$caudate2_Path_Side <- recode(demo$caudate2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$internal2_capsule_Path_Side <- recode(demo$internal_capsule2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$basal_ganigla2_Path_Side <- recode(demo$basal_ganglia2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$insula2_Path_Side <- recode(demo$insula2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$putamen2_Path_Side <- recode(demo$putamen2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$globus_pallidus2_Path_Side <- recode(demo$globus_pallidus2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$corona_radiata2_Path_Side <- recode(demo$corona_radiata2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$periventricular_white_matter2_Path_Side <- recode(demo$periventricular_white_matter2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$subcortical_white_matter2_Path_Side <- recode(demo$subcortical_white_matter2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

# 12.1.2019; Did not run because no data here
# 01.09.2020; Did not run because no data here
# 01.24.2020; Did not run because no data here
# 08.22.2020; Did not run because no data here
demo$cingulate_gyrus2_Path_Side <- recode(demo$cingulate_gyrus2_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

demo$hippocampus_secondary_Path_Side <- recode(demo$hippocampus_secondary_Path_Side, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

# Change column name
demo <- demo %>% rename("hippocampus2_Path_Side" = "hippocampus_secondary_Path_Side")
```

# EVD/ICP Monitors
```{r}
# evd: 0, No | 1, Bolt/ICP Monitor | 2, EVD
demo <- demo %>% rename(ICP_EVD_Monitor_No = evd___0, 
                        ICP_Monitor = evd___1, 
                        EVD = evd___2)
```

# Surgery

*Note: We have a separate surgery file we will use*

```{r}
# 0, None | 1, Craniectomy | 2, Craniotomy | 3, Lobectomy | 4, SDH Evacuation | 5, EDH Evacuation | 6, Ventriculoperitoneal Shunt | 7, Thrombectomy | 8, Cranioplasty | 999, Other Brain Surgery
demo <- demo %>% rename(Surgery_No = surgery___0, 
                        Craniectomy = surgery___1, 
                        Craniotomy = surgery___2, 
                        Lobectomy = surgery___3,
                        SDH_Evac = surgery___4,
                        EDH_Evac = surgery___5,
                        V_Shunt = surgery___6,
                        Thrombectomy = surgery___7,
                        Cranioplasty = surgery___8,
                        Other_Surgery = surgery___999)
                           
# Other surgery note
demo <- demo %>% rename(other_desc_surgery = other_surgery)
```                  

# Mechanical ventilation and eye problems
```{r}
# mech_vent: 1, Yes | 0, No
demo$mech_vent <- recode(demo$mech_vent, '1' = "Yes", '0' = "No")

#eye_problems:1, Cataract | 2, Glaucoma | 3, Retinal Detachment | 0, No | 999, Other
demo <- demo %>% rename(Eye_Cataract = eye_problems___1, 
                        Eye_Glaucoma = eye_problems___2, 
                        Eye_Retinal_Detachment = eye_problems___3, 
                        Eye_No_Prob = eye_problems___0,
                        Eye_Other = eye_problems___999)
                             
# cataract: 1, Left | 2, Right | 3, Both | 999, Unspecified
demo$cataract <- recode(demo$cataract, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

# glaucoma: 1, Left | 2, Right | 3, Both | 999, Unspecified 
demo$glaucoma <- recode(demo$glaucoma, '1' = "Left", '2' = "Right", '3' = "Both", '999' = "Unspecified")

# Other eye problems notes - eye Side
demo <- demo %>% rename(other_eye_prob_side = other_eye_problem)

# Other eye problems notes
demo <- demo %>% rename(other_desc_eye_prob = other_eye_problems)
                            
# The following surgery columns I'm not touching for now since i did not touch this instrument 
#"surgeries_type"                         
#"other_surgery_time"                     
#"date_surgeries"                         
#"surgeries_complete"          
```

# TBI and aSAH Outcomes
```{r}
#final_diagnosis: 1, TBI | 2, aSAH
demo$final_diagnosis <- recode(demo$final_diagnosis, '1' = "TBI", '2' = "aSAH") 

# neuro_exam_arrival notes
demo <- demo %>% rename(neuro_exam_desc = neuro_exam_arrival)

# gcs eye arrival: 1, 1 No eye opening | 2, 2 Open to pain | 3, 3 Open to verbal command | 4, 4 Open spontaneously | 999, Unable to determine
demo$gcs_eye_arrival <- recode(demo$gcs_eye_arrival, '1' = "No Eye Opening", '2' = "Open to Pain", '3' = "Open to Verbal Command", '4' = "Open Spontaneously", '999' = "Unable to Determine") 
                     
#gsc_verbal_arrival: 1, 1 No verbal response | 2, 2 Incomprehensible sounds | 3, 3 Inappropriate words | 4, 4 Confused | 5, 5 Oriented | 999, Unable to determine
demo$gsc_verbal_arrival <- recode(demo$gsc_verbal_arrival, '1' = "No Verbal Response", '2' = "Incomprehensible Sounds", '3' = "Inappropriate Words", '4' = "Confused", '5' = "Oriented", '999' = "Unable to Determine")

# Change column name - spelling error
demo <- demo %>% rename(gcs_verbal_arrival = gsc_verbal_arrival)
                   
#gsc_motor_arrival: 1, 1 No motor response | 2, 2 Extension to pain | 3, 3 Flexion to pain | 4, 4 Withdrawal from the pain | 5, 5 Localizing pain | 6, 6 Obeys commands | 999, Unable to determine
demo$gsc_motor_arrival <- recode(demo$gsc_motor_arrival, '1' = "No Motor Response", '2' = "Extension to Pain", '3' = "Flexion to Pain", '4' = "Withdrawal from Pain", '5' = "Localizing Pain", '6' = "Obeys Commands", '999' = "Unable to Determine")          

# Change column name - spelling error
demo <- demo %>% rename(gcs_motor_arrival = gsc_motor_arrival)

#vent_arrival: 1, Yes | 0, No
demo$vent_arrival <- recode(demo$vent_arrival, '1' = "Yes", '0' = "No") 
                    
# gcs eye 24: 1, 1 No eye opening | 2, 2 Open to pain | 3, 3 Open to verbal command | 4, 4 Open spontaneously | 999, Unable to determine
demo$gcs_eye_24 <- recode(demo$gcs_eye_24, '1' = "No Eye Opening", '2' = "Open to Pain", '3' = "Open to Verbal Command", '4' = "Open Spontaneously", '999' = "Unable to Determine") 
                     
#gcs_verbal_24: 1, 1 No verbal response | 2, 2 Incomprehensible sounds | 3, 3 Inappropriate words | 4, 4 Confused | 5, 5 Oriented | 999, Unable to determine
demo$gcs_verbal_24 <- recode(demo$gcs_verbal_24, '1' = "No Verbal Response", '2' = "Incomprehensible Sounds", '3' = "Inappropriate Words", '4' = "Confused", '5' = "Oriented", '999' = "Unable to Determine")

#gcs_motor_24: 1, 1 No motor response | 2, 2 Extension to pain | 3, 3 Flexion to pain | 4, 4 Withdrawal from the pain | 5, 5 Localizing pain | 6, 6 Obeys commands | 999, Unable to determine
demo$gcs_motor_24 <- recode(demo$gcs_motor_24, '1' = "No Motor Response", '2' = "Extension to Pain", '3' = "Flexion to Pain", '4' = "Withdrawal from Pain", '5' = "Localizing Pain", '6' = "Obeys Commands", '999' = "Unable to Determine")     
                    
#vent_24: 1, Yes | 0, No
demo$vent_24 <- recode(demo$vent_24, '1' = "Yes", '0' = "No")                          
                        
# gcs_eye_discharge: 1, 1 No eye opening | 2, 2 Open to pain | 3, 3 Open to verbal command | 4, 4 Open spontaneously | 999, Unable to determine
demo$gcs_eye_discharge <- recode(demo$gcs_eye_discharge, '1' = "No Eye Opening", '2' = "Open to Pain", '3' = "Open to Verbal Command", '4' = "Open Spontaneously", '999' = "Unable to Determine") 
                     
#gsc_verbal_discharge: 1, 1 No verbal response | 2, 2 Incomprehensible sounds | 3, 3 Inappropriate words | 4, 4 Confused | 5, 5 Oriented | 999, Unable to determine
demo$gcs_verbal_discharge <- recode(demo$gcs_verbal_discharge, '1' = "No Verbal Response", '2' = "Incomprehensible Sounds", '3' = "Inappropriate Words", '4' = "Confused", '5' = "Oriented", '999' = "Unable to Determine")

#gcs_motor_discharge: 1, 1 No motor response | 2, 2 Extension to pain | 3, 3 Flexion to pain | 4, 4 Withdrawal from the pain | 5, 5 Localizing pain | 6, 6 Obeys commands | 999, Unable to determine
demo$gcs_motor_discharge <- recode(demo$gcs_motor_discharge, '1' = "No Motor Response", '2' = "Extension to Pain", '3' = "Flexion to Pain", '4' = "Withdrawal from Pain", '5' = "Localizing Pain", '6' = "Obeys Commands", '999' = "Unable to Determine")     
                    
#vent_discharge: 1, Yes | 0, No
demo$vent_discharge <- recode(demo$vent_discharge, '1' = "Yes", '0' = "No")    

#neurooutcomes_discharge: 0, Dead | 1, Eye opening and not following commands | 2, No following commands | 3, Following commands | 4, Unable to determine | 999, No follow up
demo$neurooutcomes_discharge <- recode(demo$neurooutcomes_discharge, '0' = "Dead", '1' = "Eye opening - not following commands", '2' = "Not following commands", '3' = "Following Commands", '4' = "Unable to Determine", '999' = "No Follow Up")   

# Three months

#neurooutcomes_3: 0, Dead | 1, Eye opening but not following commands | 2, No following commands | 3, Following commands | 4, Unable to determine | 999, No follow up
demo$neurooutcomes_3 <- recode(demo$neurooutcomes_3, '0' = "Dead", '1' = "Eye opening - not following commands", '2' = "Not following commands", '3' = "Following Commands", '4' = "Unable to Determine", '999' = "No Follow Up")     

# gcs_eye_3: 1, 1 No eye opening | 2, 2 Open to pain | 3, 3 Open to verbal command | 4, 4 Open spontaneously | 999, Unable to determine
demo$gcs_eye_3 <- recode(demo$gcs_eye_3, '1' = "No Eye Opening", '2' = "Open to Pain", '3' = "Open to Verbal Command", '4' = "Open Spontaneously", '999' = "Unable to Determine") 
                     
#gsc_verbal_3: 1, 1 No verbal response | 2, 2 Incomprehensible sounds | 3, 3 Inappropriate words | 4, 4 Confused | 5, 5 Oriented | 999, Unable to determine
demo$gcs_verbal_3 <- recode(demo$gcs_verbal_3, '1' = "No Verbal Response", '2' = "Incomprehensible Sounds", '3' = "Inappropriate Words", '4' = "Confused", '5' = "Oriented", '999' = "Unable to Determine")

#gcs_motor_3: 1, 1 No motor response | 2, 2 Extension to pain | 3, 3 Flexion to pain | 4, 4 Withdrawal from the pain | 5, 5 Localizing pain | 6, 6 Obeys commands | 999, Unable to determine
demo$gcs_motor_3 <- recode(demo$gcs_motor_3, '1' = "No Motor Response", '2' = "Extension to Pain", '3' = "Flexion to Pain", '4' = "Withdrawal from Pain", '5' = "Localizing Pain", '6' = "Obeys Commands", '999' = "Unable to Determine")       

#vent_3: 1, Yes | 0, No
demo$vent_3 <- recode(demo$vent_3, '1' = "Yes", '0' = "No")     

# Six Months
#neurooutcomes_6: 0, Dead | 1, Eye opening but not following commands | 2, No following commands | 3, Following commands | 4, Unable to determine | 999, No follow up
demo$neurooutcomes_6 <- recode(demo$neurooutcomes_6, '0' = "Dead", '1' = "Eye opening - not following commands", '2' = "Not following commands", '3' = "Following Commands", '4' = "Unable to Determine", '999' = "No Follow Up")     

# gcs_eye_6: 1, 1 No eye opening | 2, 2 Open to pain | 3, 3 Open to verbal command | 4, 4 Open spontaneously | 999, Unable to determine
demo$gcs_eye_6 <- recode(demo$gcs_eye_6, '1' = "No Eye Opening", '2' = "Open to Pain", '3' = "Open to Verbal Command", '4' = "Open Spontaneously", '999' = "Unable to Determine") 
                     
#gsc_verbal_6: 1, 1 No verbal response | 2, 2 Incomprehensible sounds | 3, 3 Inappropriate words | 4, 4 Confused | 5, 5 Oriented | 999, Unable to determine
demo$gcs_verbal_6 <- recode(demo$gcs_verbal_6, '1' = "No Verbal Response", '2' = "Incomprehensible Sounds", '3' = "Inappropriate Words", '4' = "Confused", '5' = "Oriented", '999' = "Unable to Determine")

#gcs_motor_6: 1, 1 No motor response | 2, 2 Extension to pain | 3, 3 Flexion to pain | 4, 4 Withdrawal from the pain | 5, 5 Localizing pain | 6, 6 Obeys commands | 999, Unable to determine
demo$gcs_motor_6 <- recode(demo$gcs_motor_6, '1' = "No Motor Response", '2' = "Extension to Pain", '3' = "Flexion to Pain", '4' = "Withdrawal from Pain", '5' = "Localizing Pain", '6' = "Obeys Commands", '999' = "Unable to Determine")       

#vent_6: 1, Yes | 0, No
demo$mech_vent_6 <- recode(demo$mech_vent_6, '1' = "Yes", '0' = "No")     

# Twelve Months
#neurooutcomes_12: 0, Dead | 1, Eye opening but not following commands | 2, No following commands | 3, Following commands | 4, Unable to determine | 999, No follow up
demo$neurooutcomes_12 <- recode(demo$neurooutcomes_12, '0' = "Dead", '1' = "Eye opening - not following commands", '2' = "Not following commands", '3' = "Following Commands", '4' = "Unable to Determine", '999' = "No Follow Up")     

# gcs_eye_12: 1, 1 No eye opening | 2, 2 Open to pain | 3, 3 Open to verbal command | 4, 4 Open spontaneously | 999, Unable to determine
demo$gcs_eye_12 <- recode(demo$gcs_eye_12, '1' = "No Eye Opening", '2' = "Open to Pain", '3' = "Open to Verbal Command", '4' = "Open Spontaneously", '999' = "Unable to Determine") 
                     
#gsc_verbal_12: 1, 1 No verbal response | 2, 2 Incomprehensible sounds | 3, 3 Inappropriate words | 4, 4 Confused | 5, 5 Oriented | 999, Unable to determine
demo$gcs_verbal_12 <- recode(demo$gcs_verbal_12, '1' = "No Verbal Response", '2' = "Incomprehensible Sounds", '3' = "Inappropriate Words", '4' = "Confused", '5' = "Oriented", '999' = "Unable to Determine")

#gcs_motor_12: 1, 1 No motor response | 2, 2 Extension to pain | 3, 3 Flexion to pain | 4, 4 Withdrawal from the pain | 5, 5 Localizing pain | 6, 6 Obeys commands | 999, Unable to determine
demo$gcs_motor_12 <- recode(demo$gcs_motor_12, '1' = "No Motor Response", '2' = "Extension to Pain", '3' = "Flexion to Pain", '4' = "Withdrawal from Pain", '5' = "Localizing Pain", '6' = "Obeys Commands", '999' = "Unable to Determine")       

#vent_12: 1, Yes | 0, No
demo$vent_12 <- recode(demo$vent_12, '1' = "Yes", '0' = "No")     
```

# Mass effect and diabetes processing
```{r}
# mass_effect_severity: 1, Mild | 2, Moderate | 3, Severe
demo$mass_effect_severity <- recode(demo$mass_effect_severity, '1' = "Mild", '2' = "Moderate", '3' = "Severe")  

# mass_effect_location: 1, Supratentorial left | 2, Spratentorial right | 3, Infratentorial | 4, Supratentorial bilateral
demo <- demo %>% rename(ME_location_Supratentorial_l = mass_effect_location___1, 
                        ME_location_Supratentorial_r = mass_effect_location___2, 
                        ME_location_Infratentorial = mass_effect_location___3, 
                        ME_location_Supratentorial_bilat = mass_effect_location___4)

# diabetes: 1, Yes | 0, No | 999, Indeterminate
demo$diabetes <- recode(demo$diabetes, '1' = "Yes", '0' = "No", '999' = "Indeterminate")  
```

# Add surgery Information

```{r}
surgery <- surgery %>% 
  mutate(DateSurgery = as.POSIXct(date_surgeries))

demo <- demo %>% 
  select(-date_surgeries) %>%
  left_join(., surgery %>% select(Study_ID, DateSurgery), by = "Study_ID")
```

# Convert certain variables to factors
```{r}
# There may be numeric variables - maybe gcs that will later have to be converted -- I think it would be problematic with NAs,to not convert to factor though
demo <- demo %>% 
  mutate(Study_ID = as.character(as.numeric(Study_ID)))
```

# Save Demographics
```{r}
save(demo, file = "processed_data/Demographics.rda")
```

